---
title: "Vaccine Rate Mini Project"
author: 'Natasha ( PID: A15393874)'
date: "11/26/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

Get started
```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```
   ##Ensure the data column is useful 

We will use the **lubridate** package to make life a lot easier when dealing with dates and times

```{r}
##install.packages("lubridate")
library(lubridate)
```

```{r}
today()
```
Here we make our 'as_of_date' column lubridate format 

```{r}
# Specify that we are using the Year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

Now I can do useful math with dates easily:

```{r}
today() - vax$as_of_date[1]
```
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
```{r}
today() - vax$as_of_date[nrow(vax)]
```
> Q1. What column details the total number of people fully vaccinated?

```{r}
colnames(vax)
```
> [9] "persons_fully_vaccinated

> Q2. What column details the Zip code tabulation area?
[2] "zip_code_tabulation_area" 

> Q3. What is the earliest date in this dataset? 

```{r}
min(vax$as_of_date)

```
> Q4. What is the latest date in this dataset? 

```{r}
max(vax$as_of_date) 
```
```{r}
skimr::skim(vax)
```
> Q5. How many numeric columns are in this dataset?
9
> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?
8355
> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?
10% of persons_fully_vaccinated values are missing
> Q8. [Optional]: Why might this data be missing?
People not sharing their personal information?
> Q9. How many days have passed since the last update of the dataset?

```{r}
today() - vax$as_of_date[nrow(vax)]
```
> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```




##Working with zip codes


We will use the **zipcodeR** package to help make sense of zip codes


```{r}
#install.packages("zipcodeR")
library(zipcodeR)
geocode_zip('92037')
```
Calculate the distance between the centroids of any two ZIP codes in miles, e.g.

```{r}
zip_distance('92037','92109')
```
More usefully, we can pull census data about ZIP code areas (including median household income etc.). For example:

```{r}
reverse_zipcode(c('92037', "92109") )
```
We can use this reverse_zipcode() to pull census data later on for any or all ZIP code areas we might be interested in.
```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```
#Focus on San Diego County
# Subset to San Diego county only areas

```{r}
sd <- vax$county == "San Diego"
head(vax[sd,])
```
But let's use the **dplyr** package and it's **filter()*** function:

```{r}
library(dplyr)
sd <- filter(vax, county == "San Diego")
nrow(sd)
```
```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

> Q11. How many distinct zip codes are listed for San Diego County?

```{r}
length(unique(sd$zip_code_tabulation_area))
```
> 107 distinct zip codes

> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
sd[which.max(sd$age12_plus_population),]
```
> 92154

What is the population in the 92037 ZIP code area?
```{r}
filter(sd, zip_code_tabulation_area == "92037")[1,]
```
36144

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

```{r}
mean((filter(vax, county == "San Diego" & as_of_date == "2021-11-09"))$percent_of_population_fully_vaccinated, na.rm = TRUE)
```
> 67%

```{r}
sd.now <- filter(sd, as_of_date == "2021-11-09")
mean(sd.now$percent_of_population_fully_vaccinated, na.rm = TRUE)
```


We can look at the 6-number summary
```{r}
sd.sum <- summary(sd.now$percent_of_population_fully_vaccinated)
sd.sum
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```
```{r}
library(ggplot2)
ggplot(sd.now) +
  aes(percent_of_population_fully_vaccinated)+geom_histogram(bin=15)
```
What about 92037 - UCSD/ La Jolla
```{r}
lj <- filter(sd.now, zip_code_tabulation_area == "92037")
lj
```
> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```
```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x= "Date", y="Percent Vaccinated")
```
##Time series of vaccination rate for 92037
```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```
> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```
 > mean = 66.4%
 
```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() + 
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x= "Date", y="Percent Vaccinated") +
  geom_hline(yintercept = 0.664, colour = "red", linetype = 2)
```
> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```
> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated)+geom_histogram(bin=15) + 
  labs(x= "Percent Vaccinated", y="Count")+ 
  xlim(c(0,1))+
  ylim(c(0,70))
```
> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
reverse_zipcode(c('92109', "92040") )
```
```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
> Lakeside is below the average of 66.4%. 

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
> San Diego is above the average of .664. 

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144

irst we need to subset the full vax dataset to include onl ZIP code areas with a population as large as 92037
```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)
```
How many unique zip codes have a population as large as 92037?
```{r}
length(unique(vax.36.all$zip_code_tabulation_area))
```
```{r}
ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1.00)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccinte rate across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.6640, linetype=2)
```
> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?
Really nervous because of these statistics and how places are still below average for vaccines and also because of the peak of cases we had last year at this time.
















