---
title: "R Notebook"
author: "Natasga (PID: A15393874)
output: html_document
---
First up is clustering methods

# Kmeans clustering

The function in base R to do Kmeans clustering is called `kmeans()`.

First make up some data where we know what the answer should be:

```{r}
tmp <- c(rnorm(30, -3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


> Q. Can we use kmeans() to cluster this data setting k 2 and nstart to 20?

```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

> Q. How many point are in each cluter?

```{r}
km$size
```

> Q. What 'component' of your result object details cluster assignment/membership?

```{r}
km$cluster
```


> Q. What 'component' of your result object details cluster center?

```{r}
km$centers
```


> Q. Plot x colored by the kmeans cluster assignment an add cluster centers as blue points

```{r}
plot(x, col= km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```

This an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code. 

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
