---
title: "Class 11 Pt2"
author: 'Natasha ( PID: A15393874)'
date: "11/04/2021"
output: pdf_document
---

Using the Bio3D package
```{r}
library(bio3d)
pdb <- read.pdb("1hel")
```
```{r}
pdb

```
```{r}
aa123(pdbseq(pdb))
```
```{r}
#Atom records
head(pdb$atom)
```
> Q7: How many amino acid residues are there in this pdb object?
198
> Q8: Name one of the two non-protein residues?
MK1
> Q9: How many protein chains are in this structure?
2

```{r}
attributes(pdb)
```
```{r}
head(pdb$atom)
```
> Q10. Which of the packages above is found only on BioConductor and not CRAN?
"msa"
> Q11. Which of the above packages is not found on BioConductor or CRAN?:
"Grantlab/bio3d-view" because it is the only one that doesn't use install.packages() or BiocManager::install()
> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?
True

```{r}
library(bio3d)
aa <- get.seq("1ake_A")
```
```{r}
aa
```
> Q13. How many amino acids are in this sequence, i.e. how long is this sequence?
214

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A','4PZL_A')
```

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
Multiple Structure Alignment
```{r}
#Align related PDBs
pdbs <- pdbaln(files, fit= TRUE, exefile="msa")
```
```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```
Use bio3D pca() function
```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```
```{r}
# Calculate RMSD
rd <- rmsd(pdbs)
```
```{r}
# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k= 3)
```

```{r}
plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

